---
title: "Lab 5 244"
author: "Rachel Behm"
date: "February 7, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A. Attach Packages
```{r}
library(tidyverse)
library(tseries)
library(forecast)
```

B. Get data
US Residential energy consumtion 1/1973-10/2017
```{r}
energy <- read_csv("energy.csv")
```

month and year information together and we dont want that- what do?
we only care about res_total

C. create ts timeseries data using res_total column 
```{r}
res_ts <- ts(energy$res_total, frequency=12, c(1973,1))

#$ allows us to pull the vector form of that column 

#now lets plot to look at (shows us the time series data we just made)
plot(res_ts)


#trend= yes increase then level off
#additive= yes, but intermediate peaks increasing
```

d. Decompose timeseries data
```{r}
res_dc <- decompose(res_ts)
#wow so simple does all the work woo

plot(res_dc)
#look at scales
#seasonality scale big= big contributor
```

e. cool ways to present timeseries data
```{r}
monthplot(res_ts)
#breaks down by month how values are changing for the years
#we can see huge increase during the summer
```
```{r}

ggseasonplot(res_ts)
#can see general increase by year as well as monthly change (good for checking seasons)
#similar story in diff way

ggseasonplot(res_ts, polar= TRUE)
#makes it circular

```

f. Simple moving averages
use forecase;;ma() to specify windowsize
```{r}
#if want window size to be 5 months
sma_res <- ma(res_ts, order = 5)
#nas not a suprise since we set window to 5
plot(sma_res)
# we can see quite a bit smoother

sma_res2 <- ma(res_ts, order = 12)
plot(sma_res2)

#CHANGING WINDOW DRASTICALLY ALTERS- WATCH OUT
#solution= show both MA and original


plot(res_ts)
lines(sma_res, col= "red")
#now they are plotted together! so you/readers can compare
```

g. ACF (autocorrelation)
```{r}
#acf function does all the annoying work for us

res_acg<- acf(res_ts)
#drastic and significant peaks every 12 months (makes sense: winter vs summer)
#can confirm seasonality important


#use ggtsdisplay puts everything together
ggtsdisplay(res_ts)
#shows original on top, then ACF on bottom, we dont care about pacf yet
```

1. Holt-Winters: exponential smoothing (stronger weights to recent)
```{r}
#can predict that HW will work very well since we see strong seasonality

res_hw <- HoltWinters(res_ts)
res_hw

plot(res_hw)
#ya looks pretty darn good

#now we wanna do some forecasting

```

Future forcasting~~~~~~with HW
```{r}
res_forecast <- forecast(res_hw, h = 60)
#h=60 = 5 years
plot(res_forecast)

#trend not really increasing because less weight to older



#if we wannna plot in ggplot can call whatever then transform to df then plot in ggplot

#now lets check residuals
```

Check distribution of the residuals (make sure normal)
```{r}
#use $ to call components you want
hist(res_forecast$residuals)
#ya looks pretty normal
```

ARIMA (AutoRegressive integrated moving average) using auto.arima
```{r}
#use auto.arima() for p, d, q
res_pdq <- auto.arima(res_ts)
res_pdq

```

pdq= (1,0,2)= nonseasonal (0,1,1)=seasonal
now lets fit the arima model using these inputs

```{r}
res_arima <- arima(res_ts,c(1,0,2),seasonal = list(order= c(0,1,1)))
#seasonal has to be a list idk why it wants it like that
res_arima


#now lets check the residuals
par(mfrow = c(1,2))
hist(res_arima$residuals)
qqnorm(res_arima$residuals)
#par tells it how to be displayed= 1,2 means one row two columns (why theyre in the same pic)
#looks relatively normal
```

Now ARIMA forecasting!
```{r}
forecast_res <- forecast(res_arima, h= 72)
plot(forecast_res)
#looks pretty similar to the holt winters one = as expected
```

Let's show how to ggplot it
```{r}
#we have to coerce to df
res_df <- data.frame(forecast_res)
#makes forecast_res a dataframe = woooo

#now lets make month sequence so we dont have to deal with the words (change to 1-72)
month_seq<- seq(1,72)

res_df_2 <- data.frame(month_seq, res_df)
#binds them together



ggplot(res_df_2, aes(x= month_seq, y= Point.Forecast))+ 
  geom_line()
```

